---
title: "SCdetMito Package Documentation"
vignette: >
  %\VignetteIndexEntry{SCdetMito Package Documentation}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

# SCdetMito Package Documentation

Version: 1.3.5

Updated: 2026-05-18

Author: Silu Hu

## Purpose

SCdetMito is a lightweight R package for retention-loss enrichment-based
adaptive mitochondrial ratio cutoff recommendation in single-cell RNA-seq
quality control. The package is designed for one-sample, multi-sample, and
group-aware integrated Seurat workflows.

The package should be used as a quality-control recommendation and auditing
tool. It does not claim that a single universal mitochondrial cutoff exists for
all tissues, species, chemistries, or disease contexts, and it should
not be framed as a classical breakpoint model.

## Bundled Test Data

The installed package includes two small Seurat objects:

- `load_demo_seurat()`: synthetic multi-group toy data for fast examples.
- `load_pbmc3k_demo()`: lightweight public PBMC3K subset with 300 cells and
  700 features.

The bundled objects are intended for package examples and checks, not for
large-scale validation claims.

## Core Workflow

```{r}
library(SCdetMito)

pbmc3k <- load_pbmc3k_demo()

det <- SCdetMito(
  pbmc3k,
  sample_col = "sample",
  mito_col = "mitoRatio",
  min_cut = 0.01,
  max_cut = 0.25,
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

det$cutoff
det$sample_cutoff_summary
det$retained_cell_profile
det$interval_loss_table
```

## Strategy Guide

- `consensus`: use one all-sample sample-supported global cutoff.
- `strictest`: use the most conservative sample-level cutoff.
- `groupwise`: use group-specific cutoffs when each biological group contains
  multiple samples.

Available detection tests can be listed with:

```{r}
SCdetMito_methods()
```

Primary default:

- `loss_test = "mad_zscore"`
- `p_adj_method = "BH"`
- `sample_cutoff_method = "largest_drop"`

Useful sensitivity checks:

- `loss_test = "empirical_tail"`
- `loss_test = "poisson_tail"`
- `p_adj_method = "none"` for unadjusted exploratory review
- `p_adj_method = "holm"`
- `SCdetMito_sensitivity()` for parameter stability checks

## Citation

```{r}
citation("SCdetMito")
```

The only package author is Silu Hu.
