---
title: "SCdetMito Workflow"
vignette: >
  %\VignetteIndexEntry{SCdetMito Workflow}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`SCdetMito` provides reference-aware mitochondrial QC decision support for
Seurat workflows. The package reports retained-cell profiles,
interval-specific cell loss, literature-informed reference cutoffs, and
multiple sample-level cutoff candidates.

## Load the bundled PBMC demo

```{r}
library(SCdetMito)

pbmc <- load_demo_pbmc()
table(pbmc$sample)
table(pbmc$group)
summary(pbmc$mitoRatio)
```

The bundled demo is a downsampled multi-donor PBMC object derived from public
10x Genomics GEM-X Human PBMC Donor 1-4 datasets. It is intended for software
examples and tests, not biological inference.

## Reference-aware sample cutoffs

```{r}
det <- SCdetMito(
  pbmc,
  sample_col = "sample",
  mito_col = "mitoRatio",
  species = "human",
  tissue = "PBMC",
  sample_cutoff_method = "reference_guided",
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

det$sample_cutoff_summary[, c(
  "sample",
  "reference_cutoff",
  "first_significant_cutoff_high",
  "largest_drop_cutoff",
  "selected_cutoff",
  "detected_cutoff",
  "recommended_cutoff",
  "recommendation_level",
  "warning_level"
)]
```

`selected_cutoff` is determined by `sample_cutoff_method`. `detected_cutoff`
is retained as a backward-compatible alias of `selected_cutoff`.
`recommended_cutoff` is SCdetMito's default reference-aware recommendation and
is reported even when a strict or exploratory selected mode is requested.

## Group-aware multi-sample QC

```{r}
qc_groupwise <- SCQCmulti(
  pbmc,
  sample_col = "sample",
  group_col = "group",
  mito_col = "mitoRatio",
  cutoff_strategy = "groupwise",
  scdet_options = list(sample_cutoff_method = "reference_guided"),
  species = "human",
  tissue = "PBMC",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  write_plots = FALSE,
  write_tables = FALSE
)

qc_groupwise@misc$SCdetMito_QC$group_plan
```

## Benchmarking and sensitivity

```{r}
benchmark <- SCQCbenchmark(
  pbmc,
  sample_col = "sample",
  group_col = "group",
  species = "human",
  tissue = "PBMC",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  run_downstream = FALSE,
  plot = FALSE,
  output_dir = tempdir()
)

benchmark$recommended_strategies
```

```{r}
sensitivity <- SCdetMito_sensitivity(
  pbmc,
  sample_col = "sample",
  mito_col = "mitoRatio",
  species = "human",
  tissue = "PBMC",
  sample_cutoff_method = c("largest_drop", "reference_guided"),
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

sensitivity$sensitivity_summary
```

Benchmark scores are user-adjustable summaries. They do not imply a
universally applicable QC strategy.
