---
title: "SCdetMito Detailed Tutorial"
vignette: >
  %\VignetteIndexEntry{SCdetMito Detailed Tutorial}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

This tutorial outlines an internet-independent SCdetMito workflow using the
bundled multi-donor PBMC demo.

## Installation

Install SCdetMito from GitHub with:

```r
install.packages("remotes")
remotes::install_github("ericHuu/SCdetMito@v1.4.3", build_vignettes = FALSE)
```

Then load the package:

```{r}
library(SCdetMito)
```

## Load Demo Data

```{r}
pbmc <- load_demo_pbmc()
table(pbmc$sample)
table(pbmc$group)
summary(pbmc$mitoRatio)
```

The demo is derived from public 10x Genomics GEM-X Human PBMC Donor 1-4
datasets and downsampled to 2,500 cells per donor. It is intended for software
demonstration and testing only.

## Ensure Mitochondrial Fraction Metadata

```{r}
pbmc <- ensure_mito_ratio(
  pbmc,
  mito_col = "mitoRatio",
  species = "human"
)
```

If automatic feature detection is not appropriate, provide a complete
mitochondrial feature vector retained in the object:

```{r}
mito_features <- grep("^MT-", rownames(pbmc), value = TRUE)
pbmc <- ensure_mito_ratio(
  pbmc,
  mito_col = "mitoRatio",
  species = "human",
  mito_features = mito_features,
  overwrite = TRUE
)
```

## Run SCdetMito

```{r}
det <- SCdetMito(
  pbmc,
  sample_col = "sample",
  mito_col = "mitoRatio",
  species = "human",
  tissue = "PBMC",
  sample_cutoff_method = "reference_guided",
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

det$sample_cutoff_summary
```

`selected_cutoff` follows the requested cutoff mode. `detected_cutoff` is a
backward-compatible alias of `selected_cutoff`. `recommended_cutoff` is always
reported using the default reference-aware recommendation policy, so users can
compare the selected cutoff with SCdetMito's recommendation and warning fields.

## Apply Groupwise QC

```{r}
qc <- SCQCmulti(
  pbmc,
  sample_col = "sample",
  group_col = "group",
  mito_col = "mitoRatio",
  cutoff_strategy = "groupwise",
  scdet_options = list(sample_cutoff_method = "reference_guided"),
  species = "human",
  tissue = "PBMC",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  write_plots = FALSE,
  write_tables = FALSE
)

qc@misc$SCdetMito_QC$cutoff_applied
qc@misc$SCdetMito_QC$sample_cutoff_summary
qc@misc$SCdetMito_QC$group_cutoff_summary
```

For groupwise filtering, top-level `recommended_cutoff` and `cutoff_applied`
represent the final strategy-level cutoff. Per-sample evidence remains in
`sample_cutoff_summary`, and group-level aggregation remains in
`group_cutoff_summary`.

## Compare Strategies

```{r}
bench <- SCQCbenchmark(
  pbmc,
  sample_col = "sample",
  group_col = "group",
  species = "human",
  tissue = "PBMC",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  run_downstream = FALSE,
  plot = FALSE,
  output_dir = tempdir()
)

bench$recommended_strategies
```

The recommended strategy is ranked under the selected scoring profile. It is
not a universal QC optimum.
