---
title: "SCdetMito Detailed Tutorial"
vignette: >
  %\VignetteIndexEntry{SCdetMito Detailed Tutorial}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

This tutorial gives a practical end-to-end workflow for using `SCdetMito`.
The package should be treated as a lightweight R package. Large public
datasets, cross-species validation files, and manuscript outputs should remain
outside the installed package.

## Installation

Install from a local source directory:

```{r, purl=FALSE}
install.packages("remotes")
remotes::install_local("/path/to/SCdetMito")
library(SCdetMito)
```

Install from a built source package:

```{r, purl=FALSE}
install.packages(
  "/path/to/SCdetMito_1.3.5.tar.gz",
  repos = NULL,
  type = "source"
)
library(SCdetMito)
```

## Example Data

Use the synthetic multi-group object for fast function checks:

```{r, purl=FALSE}
toy <- load_demo_seurat()
table(toy$sample)
table(toy$group)
```

Use the lightweight public PBMC3K subset for recognizable real-data examples:

```{r, purl=FALSE}
pbmc3k <- load_pbmc3k_demo()
pbmc3k
table(pbmc3k$sample)
summary(pbmc3k$mitoRatio)
```

The bundled PBMC3K object is intentionally small. It is suitable for smoke
tests and tutorials, not for full manuscript claims.

## Add or Check Mitochondrial Ratio

If your Seurat object does not have a mitochondrial ratio column:

```{r, purl=FALSE}
seu <- add_mitoRatio(
  seurat_obj = seu,
  pattern = "auto",
  column = "mitoRatio",
  scale = "fraction"
)
```

For human data, `pattern = "^MT-"` is often appropriate. For mouse data,
`pattern = "^mt-"` is often appropriate. For livestock and bird datasets,
species presets can help when mitochondrial genes are not prefixed in the
usual way.

Then validate the metadata column:

```{r, purl=FALSE}
seu <- check_seu(
  seurat_obj = seu,
  check = "mitoRatio",
  normalize_fraction = TRUE,
  must_be_numeric = TRUE
)
```

## Detect an Adaptive Mitochondrial Cutoff

```{r, purl=FALSE}
det <- SCdetMito(
  seurat_obj = pbmc3k,
  by = "sample",
  mitoRatio = "mitoRatio",
  min_cut = 0.01,
  max_cut = 0.25,
  loss_test = "mad_zscore",
  p_adj_method = "BH",
  sample_cutoff_method = "largest_drop",
  min_group_support = 0.5,
  plot = FALSE,
  table_out = TRUE,
  return_details = TRUE,
  output_dir = "SCdetMito_result"
)

det$cutoff
det$sample_cutoff_summary
det$settings
```

Interpretation:

- `det$cutoff` is the final sample-supported cutoff under the selected rule.
- `det$sample_cutoff_summary` shows each sample-level cutoff.
- `det$retained_cell_profile` shows retained cells across candidate cutoffs.
- `det$interval_loss_table` shows interval-specific cell loss, candidate
  filters, adjusted p-values, and significant intervals.
- fallback sources indicate that no statistically supported retention-loss
  interval was found for that sample.

## Choose a Detection Test

```{r, purl=FALSE}
SCdetMito_methods()
```

Primary starting point:

- `loss_test = "mad_zscore"`
- `p_adj_method = "BH"`
- `sample_cutoff_method = "largest_drop"`

Sensitivity checks:

- `loss_test = "empirical_tail"`
- `p_adj_method = "none"` for unadjusted exploratory review
- `p_adj_method = "holm"`
- `sample_cutoff_method = "largest_drop"`

`threshold_only` should mainly be used for sensitivity checks, not as a final
manuscript setting.

## Sensitivity Analysis

Use `SCdetMito_sensitivity()` to evaluate whether sample-level cutoffs remain
stable across reasonable parameter choices:

```{r, purl=FALSE}
sensitivity <- SCdetMito_sensitivity(
  seurat_obj = pbmc3k,
  sample_col = "sample",
  mito_col = "mitoRatio",
  bin_width = c(0.005, 0.01, 0.02),
  alpha = c(0.01, 0.05, 0.10),
  min_drop_fraction = c(0.005, 0.01),
  loss_test = c("mad_zscore", "empirical_tail"),
  sample_cutoff_method = c("largest_drop", "max_significant"),
  write_plots = TRUE,
  output_dir = "SCdetMito_sensitivity"
)

head(sensitivity)
```

Stable cutoffs across profiles support robustness. Highly variable cutoffs
should be reported transparently and followed by biological inspection.

## Single-Sample QC

```{r, purl=FALSE}
qc_single <- SCQCone(
  seurat_obj = pbmc3k,
  min_genes = 0,
  min_counts = 0,
  max_mito = "SCdetMito",
  removeDouble = FALSE,
  plot = FALSE,
  table_out = TRUE,
  output_dir = "SCQCone_result"
)
```

Use a fixed threshold when needed:

```{r, purl=FALSE}
qc_fixed <- SCQCone(
  seurat_obj = pbmc3k,
  max_mito = 0.10,
  removeDouble = FALSE
)
```

Numeric values between 1 and 100 are interpreted as percentages, so
`max_mito = 10` is converted to `0.10` with a warning.

## Multi-Sample and Group-Aware QC

For several samples without higher-level biological groups:

```{r, purl=FALSE}
qc_consensus <- SCQCmulti(
  seurat_obj = pbmc3k,
  by = "sample",
  mode = "all",
  cutoff_strategy = "consensus",
  max_mito = "SCdetMito",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  plot = FALSE,
  table_out = TRUE,
  output_dir = "SCQCmulti_consensus"
)
```

For integrated studies with biological groups:

```{r, purl=FALSE}
qc_groupwise <- SCQCmulti(
  seurat_obj = seu,
  by = "sample_id",
  group_by = "condition",
  mode = "all",
  cutoff_strategy = "groupwise",
  max_mito = "SCdetMito",
  cutoff_support_fraction = 0.5,
  removeDouble = FALSE,
  plot = FALSE,
  table_out = TRUE,
  output_dir = "SCQCmulti_groupwise"
)

qc_groupwise@misc$SCdetMito_QC$sample_plan
qc_groupwise@misc$SCdetMito_QC$group_plan
```

Strategy guide:

- `consensus`: one all-sample sample-supported global threshold.
- `strictest`: minimum sample-level cutoff; useful when you need maximum
  stringency.
- `groupwise`: group-specific sample-supported thresholds; useful for
  multi-group studies where each group has multiple samples.

## Benchmark Fixed and Adaptive Strategies

```{r, purl=FALSE}
benchmark <- SCQCbenchmark(
  seurat_obj = pbmc3k,
  sample_col = "sample",
  strategies = list(
    list(name = "fixed_5", mode = "all", cutoff_strategy = "consensus", max_mito = 0.05),
    list(name = "fixed_10", mode = "all", cutoff_strategy = "consensus", max_mito = 0.10),
    list(name = "SCdetMito_consensus", mode = "all", cutoff_strategy = "consensus", max_mito = "SCdetMito")
  ),
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  run_downstream = FALSE,
  plot = FALSE,
  output_dir = "benchmark_result"
)

benchmark$summary
benchmark$scores
benchmark$recommended_strategies
benchmark$recommendation
```

Do not claim a method is better only because it keeps more cells. For a
publishable comparison, report retention, residual mitochondrial burden,
sample balance, group balance, cell type composition, marker stability, and
downstream clustering quality.

## Downstream Analysis

After SCdetMito QC, continue with the standard Seurat workflow:

```{r, purl=FALSE}
qc_obj <- NormalizeData(qc_obj)
qc_obj <- FindVariableFeatures(qc_obj)
qc_obj <- ScaleData(qc_obj)
qc_obj <- RunPCA(qc_obj)
qc_obj <- FindNeighbors(qc_obj, dims = 1:20)
qc_obj <- FindClusters(qc_obj, resolution = 0.5)
qc_obj <- RunUMAP(qc_obj, dims = 1:20)
```

For manuscripts, compare SCdetMito against fixed thresholds such as 5%, 10%,
15%, and 20%. Keep public validation data and large result tables outside the
R package.
