---
title: "SCdetMito Workflow"
vignette: >
  %\VignetteIndexEntry{SCdetMito Workflow}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`SCdetMito` focuses on single-cell RNA-seq quality control driven by
mitochondrial burden. The package supports both single-sample QC and
multi-sample integrated studies in which several biological groups contain
multiple samples.

## Load the demo object

```{r}
library(SCdetMito)

seu <- load_demo_seurat()
seu
table(seu$group)
table(seu$sample)
```

The bundled object is intentionally small so that the package can be tested and
demonstrated quickly. It is a smoke-test object, not a large validation
dataset. Optional public data loaders are available for explicit user-run
examples: `load_pbmc3k_online()` for a small public smoke test,
`load_heart10k_online()` for a larger real-tissue example, and
`load_ifnb_online()` for a multi-condition workflow example. These loaders are
not used during package checks.

## Single-sample QC

```{r}
sample_s1 <- subset(seu, subset = sample == "S1")

qc_single <- SCQCone(
  seurat_obj = sample_s1,
  min_genes = 0,
  min_counts = 0,
  max_mito = 0.10,
  removeDouble = FALSE,
  plot = FALSE,
  table_out = FALSE
)

ncol(sample_s1)
ncol(qc_single)
```

`SCQCone()` is appropriate when the dataset is a single sample or when you want
to apply a manually chosen threshold to one sample at a time.

## Group-aware multi-sample QC

For integrated analysis, `SCQCmulti()` can estimate sample-level mitochondrial
cutoffs from retained-cell and interval-loss profiles, then combine them
according to the selected sample-support strategy.

```{r}
qc_groupwise <- SCQCmulti(
  seurat_obj = seu,
  by = "sample",
  group_by = "group",
  mode = "all",
  cutoff_strategy = "groupwise",
  min_genes = 0,
  min_counts = 0,
  max_mito = "SCdetMito",
  removeDouble = FALSE,
  plot = FALSE,
  table_out = FALSE
)

ncol(seu)
ncol(qc_groupwise)
qc_groupwise@misc$SCdetMito_QC$group_plan
```

In practice:

- `consensus` gives one all-sample sample-supported global cutoff for the full object
- `strictest` takes the minimum sample-level cutoff
- `groupwise` allows group-specific cutoffs and is often a practical starting
  point for multi-group studies

## Benchmark QC strategies

`SCQCbenchmark()` compares fixed thresholds and adaptive strategies, then writes
summary tables and a markdown report.

```{r}
benchmark <- SCQCbenchmark(
  seurat_obj = seu,
  sample_col = "sample",
  group_col = "group",
  strategies = list(
    list(name = "fixed_10", mode = "all", cutoff_strategy = "consensus", max_mito = 0.10),
    list(name = "fixed_15", mode = "all", cutoff_strategy = "consensus", max_mito = 0.15),
    list(name = "SCdetMito_groupwise", mode = "all", cutoff_strategy = "groupwise", max_mito = "SCdetMito")
  ),
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  run_downstream = FALSE,
  plot = FALSE,
  save_objects = FALSE,
  output_dir = tempdir()
)

benchmark$recommendation
benchmark$recommended_strategies
benchmark$scores[, c("strategy", "overall_score", "retention_rate", "median_mito_after")]
```

The returned score table is intended to support method comparison and
manuscript preparation. It does not replace biological judgment, but it makes
the QC choice explicit and reproducible under the selected scoring profile.

## Recommended use in real studies

1. Compute `mitoRatio` with `add_mitoRatio()` if needed.
2. Run `SCdetMito()` or `SCQCmulti()` to estimate adaptive cutoffs.
3. Prefer `groupwise` QC when several samples exist inside each biological
   group.
4. Use `SCQCbenchmark()` to compare the adaptive strategy against fixed
   thresholds before finalizing downstream analysis.
