---
title: "SCdetMito Package Documentation"
vignette: >
  %\VignetteIndexEntry{SCdetMito Package Documentation}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

# SCdetMito Package Documentation

Version: 1.4.1

Updated: 2026-05-21

Author: Silu Hu

## Purpose

SCdetMito is an R package for reference-aware and sample-aware adaptive
mitochondrial QC in single-cell RNA-seq data. It reports literature-informed
reference cutoffs, first significant high-to-low boundaries, largest-drop
boundaries, selected cutoffs, warning fields, and provenance metadata.

The package is a QC decision-support framework. It does not define a universal
biological mitochondrial cutoff.

## Demo Data

The installed package includes one practical bundled demo:

- `load_demo_pbmc()`: a downsampled multi-donor PBMC Seurat object derived from
  public 10x Genomics GEM-X Human PBMC Donor 1-4 datasets.
- `load_demo_seurat()`: backward-compatible alias for `load_demo_pbmc()`.

The demo contains four donor-level samples, two demographic groups, raw counts,
retained mitochondrial features, and `mitoRatio` calculated with
`ensure_mito_ratio(species = "human")`. It is intended for examples and
software tests, not biological inference.

Optional online public-data loaders are retained as user-run utilities. They
are not used by standard examples, vignettes, tests, or R CMD check.

## Core Workflow

```{r}
library(SCdetMito)

pbmc <- load_demo_pbmc()

det <- SCdetMito(
  pbmc,
  sample_col = "sample",
  mito_col = "mitoRatio",
  species = "human",
  tissue = "PBMC",
  sample_cutoff_method = "reference_guided",
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

det$sample_cutoff_summary
```

## Mitochondrial Ratio Handling

```{r}
pbmc <- ensure_mito_ratio(pbmc, mito_col = "mitoRatio", species = "human")
```

Manual mitochondrial feature lists should include all mitochondrial features
retained in the object:

```{r}
mito_features <- grep("^MT-", rownames(pbmc), value = TRUE)
pbmc <- ensure_mito_ratio(pbmc, mito_features = mito_features, species = "human")
```

Short illustrative mitochondrial gene subsets should not be used for real QC.

## Multi-Sample QC

```{r}
qc_groupwise <- SCQCmulti(
  pbmc,
  sample_col = "sample",
  group_col = "group",
  mito_col = "mitoRatio",
  cutoff_strategy = "groupwise",
  scdet_options = list(sample_cutoff_method = "reference_guided"),
  species = "human",
  tissue = "PBMC",
  min_genes = 0,
  min_counts = 0,
  removeDouble = FALSE,
  write_plots = FALSE,
  write_tables = FALSE
)
```

## Reference Cutoffs

```{r}
SCdetMito_reference_cutoffs(species = "human", tissue = "PBMC")
```

Reference cutoffs are literature-informed priors for interpretation. They are
not hard filtering rules and should not be treated as tissue-invariant
thresholds.
