---
title: "SCdetMito Package Documentation"
vignette: >
  %\VignetteIndexEntry{SCdetMito Package Documentation}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

# SCdetMito Package Documentation

Version: 1.4.0

Updated: 2026-05-20

Author: Silu Hu

## Purpose

SCdetMito is a lightweight R package for reference-aware retention-loss
enrichment-based mitochondrial ratio cutoff recommendation in single-cell
RNA-seq quality control. The package reports literature-informed references,
first significant high-to-low boundaries, largest-drop boundaries, and the
selected cutoff under the requested decision mode.

The package should be used as a quality-control recommendation and auditing
tool. It does not claim that a single universal mitochondrial cutoff exists for
all tissues, species, chemistries, or disease contexts, and it should
not be framed as a classical breakpoint model.

## Example Data

The installed package includes lightweight objects for examples and checks:

- `load_demo_seurat()`: synthetic multi-group toy data for fast examples.
- `load_pbmc3k_demo()`: lightweight public PBMC3K subset with 300 cells and
  700 features.

The bundled objects are intended for package examples and checks, not for
large-scale validation claims.

For a real public dataset, users can optionally call
`load_pbmc3k_online(method = "10x")`, `load_mousebrain5k_online()`,
`load_heart10k_online()`, or `load_ifnb_online()`. PBMC3K is a small public
smoke-test example, mouse brain 5k is a tissue/nuclei scouting candidate,
heart10k is a larger solid-tissue stress-test candidate, and IFNB is a
multi-condition workflow example. Online datasets are downloaded or loaded only
when the corresponding function is called. Internet access is not required for
standard package checks.

## Core Workflow

```{r}
library(SCdetMito)

pbmc3k <- load_pbmc3k_demo()

det <- SCdetMito(
  pbmc3k,
  sample_col = "sample",
  mito_col = "mitoRatio",
  min_cut = 0.01,
  max_cut = 0.25,
  species = "human",
  tissue = "PBMC",
  sample_cutoff_method = "reference_guided",
  return_details = TRUE,
  write_plots = FALSE,
  write_tables = FALSE
)

det$cutoff
det$sample_cutoff_summary[, c(
  "sample",
  "reference_cutoff",
  "first_significant_cutoff_high",
  "largest_drop_cutoff",
  "selected_cutoff",
  "detected_cutoff",
  "warning_level"
)]
det$retained_cell_profile
det$interval_loss_table
```

`selected_cutoff` is the reference-aware decision field introduced in v1.4.0.
`detected_cutoff` is retained as a backward-compatible alias. Reference cutoffs
are literature-informed priors, not hard filtering rules; tissue-note rows may
use species-level priors unless explicitly marked as tissue-specific.

## Strategy Guide

- `consensus`: use one all-sample sample-supported global cutoff.
- `strictest`: use the most conservative sample-level cutoff.
- `groupwise`: use group-specific cutoffs when each biological group contains
  multiple samples.

Available detection tests can be listed with:

```{r}
SCdetMito_methods()
```

Primary default and SoftwareX reference-aware mode:

- `loss_test = "mad_zscore"`
- `p_adj_method = "BH"`
- `sample_cutoff_method = "largest_drop"` is retained as the default for
  backward compatibility
- `sample_cutoff_method = "reference_guided"` prioritizes the first significant
  high-to-low boundary when available and reports reference-deviation warnings

Useful sensitivity checks:

- `loss_test = "empirical_tail"`
- `loss_test = "poisson_tail"`
- `p_adj_method = "none"` for unadjusted exploratory review
- `p_adj_method = "holm"`
- `SCdetMito_sensitivity()` for parameter stability checks

## Citation

```{r}
citation("SCdetMito")
```

The only package author is Silu Hu.
